---
title: "z-test"
author: "C. Durso"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
library(openintro)
```

## Visualize areas under the standard Normal density


```{r}
x<-seq(-3,3,0.001)
dat<-data.frame(x=x,y=dnorm(seq(-3,3,0.001)))

## setting our a-value
## the positive value that has probability .975 that a standard normal random variable sampled will lie to the left of THIS value
a<-qnorm(.975) 

lhs<-data.frame(x=c(x[x<(-a)],-a,-3),y=c(dnorm(x[x<(-a)]),0,0))
rhs<-data.frame(x=c(x[x>a],3,a),y=c(dnorm(x[x>a]),0,0))

ggplot(rhs,aes(x=x,y=y))+geom_polygon(fill="gray")

g<-ggplot(dat,aes(x=x,y=y))+geom_line()+labs(title="Standard Normal Distribution")
g<-g+geom_polygon(data=lhs,aes(x=x,y=y),fill="gray",color="black")
g<-g+geom_polygon(data=rhs,aes(x=x,y=y),fill="gray",color="black")
g<-g+annotate("text",label=c("0.025","0.95","0.025"),x=c(-a-.2,0,a+.2),y=.02)
g

```

## z test

Used for (nearly) normal data with known variance, say from known measurement error. In the following, we simulate data with unknown mean m and known variance equal to 9.

### Sample a Normal distribution

Generate an unknown mean m. Make a sample from Normal(m,3), without peeking at m.

```{r}
set.seed(17)
## generate a random value between 0 and 10 for a uniform dist then take the mean
m<-mean(runif(15,0,10))
n<-6
sigma<-3
x<-rnorm(n,m,sigma)
mean(x)
```

### Is m=5?

Do a z-test of the null hypothesis that m=5.

```{r}
## 5 is the hypothesized mean; x is the sample mean
## calculate the z-score
(z<-(mean(x)-5)/(sigma/sqrt(length(x))))
## compute the 2-sided p-value
(p<-2*pnorm(-abs(z))) ## the probability of getting a z-value as extreme or more extreme than .48 (less than -.48 or greater than .48)
(p2<-2*(pnorm(-abs(mean(x)-5),0,sigma/sqrt(n))))
```

### (100-pct*100) conf. interval for m:

```{r}
pct<-.05
a<-qnorm(1-(pct)/2)
(interval<-c(mean(x)-sigma/sqrt(n)*a,mean(x)+sigma/sqrt(n)*a))
m
```

### Demo of confidence interval concept:

The (1-p)*100% confidence interval is calculated from the data using a method that has a probability of 1-p of including the true mean. Check this by drawing many samples, computing the confidence intervals, and finding the proportion of confidence intervals that include the true mean.

```{r}
set.seed(12345)
samp<-matrix(rnorm(6*1000,m,sigma),ncol=6)
conf_interval<-function(x,stdev,alpha){
  n<-length(x)
	return(c(mean(x)-stdev/sqrt(n)*alpha,mean(x)+stdev/sqrt(n)*alpha))
}
intervals<-apply(samp,1,conf_interval,stdev=sigma,alpha=a)
mean(intervals[1,]<=m & intervals[2,]>=m)
```

